Discrete Quantum Theories and Computing

### 2018-12-01 dissertation.tex
dissertation.lyx Add some TODOs for new ideas...

\begin{table}
\noindent \centering{}\caption{\label{tab:Comparison}Comparison between the classical computational
models and their physical realizations \todo{Update this table to
include }}
\begin{tabularx}{\linewidth}{ccX}
\toprule 
\addlinespace
 & Model & Physical Realization\tabularnewline\addlinespace
\midrule
\addlinespace
\addlinespace
Continuum & BCSS machine & The precision of an analog computer was limited so that an analog computer may not reproduce the theoretical prediction reliably.\tabularnewline\addlinespace
\addlinespace
\addlinespace
Discrete & Turing machine & A digital computer can be reliably repeated with exact equivalence.\tabularnewline\addlinespace
\bottomrule
\addlinespace
\end{tabularx}
\end{table}
The situations for classical computation is summarized in Table~\ref{tab:Comparison},
and similar situations could potentially be applied to quantum computation.
\todo{Introduction cleaning up until here. }Similar to classical
analog computers manipulating, say, the position of substance, the
most widely used model of quantum computing, the quantum circuit model,
manipulates the probability amplitudes to do the computation. \todo{Play
with some cloud quantum computers to see if I can find more support
evidence. }Because we are agnostic about whether probability amplitude
is quantized or can really be valued in the continuum of complex numbers,
we tried to develop two different types of models: the quantum theories
and computing over finite fields are discrete models, and the quantum
interval-valued probability measures incorporate the idea of measurement
precision into the quantum theory. To distinguish the modified quantum
theories from the usual ones, the usual ones will be called the ‘conventional
quantum theory’ where necessary and it will be reviewed in Chapter~\ref{chap:Conventional-Quantum-Theory}~\footnote{Alternative terminology in the literature includes ‘actual,’ ‘standard,’
and ‘ordinary’ quantum theory.}.

@@

Given a Hilbert space $\mathcal{H}$ of dimension $D$ and a probability
assignment for every projector $P$, we can define the expectation
value of an observable~$\mathbf{O}$ having spectral decomposition
$\mathbf{O}=\sum_{i=0}^{D-1}\lambda_{i}P_{i}$, with eigenvalues $\lambda_{i}\in\mathbb{R}$,
as \cite{544199,Jaeger2007}: 
\begin{equation}
\expval{\mathbf{O}}_{\mu}=\sum_{i=0}^{D-1}\lambda_{i}\mu\left(P_{i}\right)\,,\label{eq:quantum-expectation}
\end{equation}
where the subscript $\mu$ might be omitted if it is clear according
to the context. This definition is also consistent with the classical
expectation values because we can pullback an observable to a classical
random variable, and the expectation values are invariant.

@@

Iterating the similar process with $\ket{\Phi''}=U''\ket{\Phi'}=\frac{2\alpha}{\gamma}\left(\ket{0}+\ket{1}\right)+\frac{\beta}{\gamma}\ket{2}$,
where $\beta=\fnorm*{-1}\in\ff{p^{2}}$ and
\begin{equation}
U''=\begin{pmatrix}1 & 0 & 0\\
0 & \beta^{*} & 2\alpha\\
0 & 2\alpha^{*} & \beta
\end{pmatrix}\,.
\end{equation}
Because $\ket{\Phi''}$ is invariant under $U_{1}$, invoking Eq.~(\ref{eq:basis-independent})
in $\muF_{\Phi''}\left(P_{1}\right)=\muF_{U_{1}\ket{\Phi''}}\left(U_{1}P_{1}U_{1}^{\dagger}\right)=\muF_{\Phi''}\left(P_{0}\right)$,
we know $\muF_{\Phi''}\left(P_{1}\right)$ and $\muF_{\Phi''}\left(P_{0}\right)$
are same. Actually, they are both equal to $\tfrac{2}{3}$ because
$\muF_{\Phi''}\left(P_{0}\right)=\muF_{U''\ket{\Phi'}}\left(U''P_{0}U''^{\dagger}\right)=\muF_{\Phi'}\left(P_{0}\right)=\tfrac{2}{3}$
by Eqs.~(\ref{eq:basis-independent}) and (\ref{eq:prob23}). Since
$P'\ket{\Phi''}=\ket{\Phi''}$, Eq.~(\ref{eq:zero}) implies
\begin{equation}
1=\muF_{\Phi''}\left(P'\right)=\muF_{\Phi''}\left(P_{0}\right)+\muF_{\Phi''}\left(P_{1}\right)+\muF_{\Phi''}\left(P_{2}\right)=\tfrac{2}{3}+\tfrac{2}{3}+\muF_{\Phi''}\left(P_{2}\right)>\tfrac{4}{3}\,.
\end{equation}
This is inconsistent with the requirement that the probabilities for
orthogonal outcomes add up to $1$, and build a contradiction.\end{proof}

@@

To summarize, a conventional quantum probability measure has an uncountable
range $[0,1]$. A QIVPM constructed by blurring such a conventional
quantum probability measure must also have an uncountable range of
intervals. Of course, any particular QIVPM, or any particular experiment,
will use a fixed collection of intervals appropriate for the resources
and precision of the particular experiment.\todo{Add a two-page conclusion
chapter which could include further discussion, but it can be only
a few sentences? }\todo{Discrete Schrödinger equation? And its relation
with Grover's algorithm?}

>>>>

\begin{table}
\noindent \centering{}\caption{\label{tab:Comparison}Comparison between the classical computational
models and their physical realizations \todo{Update this table to
include }}
\begin{tabularx}{\linewidth}{ccX}
\toprule 
\addlinespace
 & Model & Physical Realization\tabularnewline\addlinespace
\midrule
\addlinespace
\addlinespace
Continuum & BCSS machine & The precision of an analog computer was limited so that an analog computer may not reproduce the theoretical prediction reliably.\tabularnewline\addlinespace
\addlinespace
\addlinespace
Discrete & Turing machine & A digital computer can be reliably repeated with exact equivalence.\tabularnewline\addlinespace
\bottomrule
\addlinespace
\end{tabularx}
\end{table}
The situations for classical computation is summarized in Table~\ref{tab:Comparison},
and similar situations could potentially be applied to quantum computation.
\todo{Introduction cleaning up until here. }Similar to classical
analog computers manipulating, say, the position of substance, the
most widely used model of quantum computing, the quantum circuit model,
manipulates the probability amplitudes to do the computation. \todo{Play
with some cloud quantum computers to see if I can find more support
evidence. }Because we are agnostic about whether probability amplitude
is quantized or can really be valued in the continuum of complex numbers,
we tried to develop two different types of models: the quantum theories
and computing over finite fields are discrete models, and the quantum
interval-valued probability measures incorporate the idea of measurement
precision into the quantum theory. To distinguish the modified quantum
theories from the usual ones, the usual ones will be called the “conventional
quantum theory” where necessary and it will be reviewed in Chapter~\ref{chap:Conventional-Quantum-Theory}~\footnote{Alternative terminology in the literature includes “actual,”
“standard,” and “ordinary” quantum theory.}.

@@

Given a Hilbert space $\mathcal{H}$ of dimension $D$ and a probability
assignment for every projector $P$, we can define the expectation
value of an observable~$\mathbf{O}$ having spectral decomposition
$\mathbf{O}=\sum_{i=0}^{D-1}\lambda_{i}P_{i}$, with eigenvalues $\lambda_{i}\in\mathbb{R}$,
as \cite{544199,Jaeger2007}: 
\begin{equation}
\expval{\mathbf{O}}_{\mu}=\sum_{i=0}^{D-1}\lambda_{i}\mu\left(P_{i}\right)\,,\label{eq:quantum-expectation}
\end{equation}
where the subscript $\mu$ might be omitted if it is clear according
to the context. This definition is also consistent with the classical
expectation values because we can pullback an observable to a classical
random variable, and the expectation values are invariant.\todo{Add
interpretation!}

@@

Iterating the similar process with $\ket{\Phi''}=U''\ket{\Phi'}=\frac{2\alpha}{\gamma}\left(\ket{0}+\ket{1}\right)+\frac{\beta}{\gamma}\ket{2}$,
where $\beta=\fnorm*{-1}\in\ff{p^{2}}$ and
\begin{equation}
U''=\begin{pmatrix}1 & 0 & 0\\
0 & \beta^{*} & 2\alpha\\
0 & 2\alpha^{*} & \beta
\end{pmatrix}\,.
\end{equation}
Because $\ket{\Phi''}$ is invariant under $U_{1}$, invoking Eq.~(\ref{eq:basis-independent})
in $\muF_{\Phi''}\left(P_{1}\right)=\muF_{U_{1}\ket{\Phi''}}\left(U_{1}P_{1}U_{1}^{\dagger}\right)=\muF_{\Phi''}\left(P_{0}\right)$,
we know $\muF_{\Phi''}\left(P_{1}\right)$ and $\muF_{\Phi''}\left(P_{0}\right)$
are same. Actually, they are both equal to $\tfrac{2}{3}$ because
$\muF_{\Phi''}\left(P_{0}\right)=\muF_{U''\ket{\Phi'}}\left(U''P_{0}U''\right)=\muF_{\Phi'}\left(P_{0}\right)=\tfrac{2}{3}$
by Eqs.~(\ref{eq:basis-independent}) and (\ref{eq:prob23}). Since
$P'\ket{\Phi''}=\ket{\Phi''}$, Eq.~(\ref{eq:zero}) implies
\begin{equation}
1=\muF_{\Phi''}\left(P'\right)=\muF_{\Phi''}\left(P_{0}\right)+\muF_{\Phi''}\left(P_{1}\right)+\muF_{\Phi''}\left(P_{2}\right)=\tfrac{2}{3}+\tfrac{2}{3}+\muF_{\Phi''}\left(P_{2}\right)>\tfrac{4}{3}\,.
\end{equation}
This is inconsistent with the requirement that the probabilities for
orthogonal outcomes add up to $1$, and build a contradiction.\end{proof}

@@

To summarize, a conventional quantum probability measure has an uncountable
range $[0,1]$. A QIVPM constructed by blurring such a conventional
quantum probability measure must also have an uncountable range of
intervals. Of course, any particular QIVPM, or any particular experiment,
will use a fixed collection of intervals appropriate for the resources
and precision of the particular experiment.\todo{Add a two-page conclusion
chapter which could include further discussion, but it can be only
a few sentences? }\todo{Discrete Schrödinger equation? And its relation
with Grover's algorithm? }\todo{Notice that in differential geometry,
Killing–Hopf theorem asserts that having positive constant curvature
guarantees a manifold is essentially a sphere, but say little things
even if the curvature has only small deviation from constant. Since
a quantum probability measure is glued by classical probability measures,
like a manifold. When their relations are exact, Gleason's theorem
asserts that a quantum probability measure can essentially induced
by the Born rule, but they can be wild when we move on to interval-valued
situation! }
